---
title: "Week 11 Lab Assignment 2"
output: 
  html_document:
    df_print: paged
author: "Jonathan Bonaguro"
date: "April 10, 2020"
---

```{r setup environment, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(collapse = TRUE)
library(tidyverse)
library(modelr)
library(broom)

mtcars <- as_tibble(mtcars)
mtcars <- mtcars %>% 
  mutate(cyl = factor(cyl, labels = c("4 Cylinders", "6 Cylinders", "8 Cylinders")))
```

## Modeling with `modelr`, `purrr`, and `broom`

You can refer to the following reference pages for questions below:  
[`tidyr`](https://tidyr.tidyverse.org/reference/nest.html)
[`purrr`](https://broom.tidyverse.org/index.html)
[`broom`](https://broom.tidyverse.org/reference/tidy.lm.html)

Use the `mtcars` dataset and answer the following questions,

1. Group this dataset by the cylinder variable.

2. Nest this grouped dataset.

3. Using the nested data, with appropriate `dplyr` and `map` functions, fit a model to predict miles per galon using weight of the cars.

4. Using `modelr` and appropriate `map` functions now add the predictions and the residuals for this nested data grouped by cylinders.

5. Use the `tidy()` and `glance()` function in the `broom` package in combination with the `map` functions to create summary statistical measures associated with your fitted model.

6. Unnest the above data and visualize the residuals.

7. Unnest the above data and present your statistical summaries across each rows of observation.

#### Assignment 1

1. Group this dataset by the cylinder variable.

```{r}
mtcars_group <- mtcars %>% 
  group_by(cyl)

mtcars_group
```

#### Assignment 2

2. Nest this grouped dataset.

```{r}
mtcars_nest <- mtcars_group %>% 
  nest()

mtcars_nest
```

#### Assignment 3

3. Using the nested data, with appropriate dplyr and map functions, fit a model to predict miles per galon using weight of the cars.

```{r}
mtcars_fit_model <- function(df) {
  lm(mpg ~ wt, data=df)
}

mtcars_model <- mtcars_nest %>% 
  mutate(model = map(data, mtcars_fit_model))
```

#### Assignment 4

4. Using modelr and appropriate map functions now add the predictions and the residuals for this nested data grouped by cylinders.

```{r}
mtcars_model_info <- mtcars_model %>% 
  mutate(residuals = map2(data, model, add_residuals)) %>% 
  mutate(predictions = map2(data, model, add_predictions))
```

#### Assignment 5

5. Use the tidy() and glance() function in the broom package in combination with the map functions to create summary statistical measures associated with your fitted model.

```{r}
mtcars_model_full_info <- mtcars_model_info %>% 
  mutate(tidy = map(model, tidy)) %>% 
  mutate(glance = map(model, glance))
```

#### Assignment 6

6. Unnest the above data and visualize the residuals.

```{r}
mtcars_model_full_info %>% 
  unnest(residuals) %>% 
  ggplot(aes(x = wt, y = resid)) +
  geom_line(aes(color = cyl)) +
  labs(title = "Residual Plot group by Number of Cylinders", subtitle = "Model: Predict Miles per Gallon by Car Weight", x = "Weight (1000 lbs.)", y= "Residuals", color = "Number of Cylinders", caption = "Motor Trends Data Provided by R")
```

#### Assignment 7

7. Unnest the above data and present your statistical summaries across each rows of observation.

```{r}
mtcars_model_full_info %>% 
  unnest(tidy) %>% 
  unnest(glance, names_repair = "unique") %>% 
  select(-(data:predictions))
```
