---
title: "Week 11 Lab Assignment 1"
output: 
  html_document:
    df_print: paged
author: "Jonathan Bonaguro"
date: "April 8, 2020"
---

```{r setup environment, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(collapse = TRUE)
library(tidyverse)
library(modelr)
library(magrittr)
```

## Modeling with `modelr`

For answering the following questions, create two continous variables `x` and `y` as per your choice using `runif()` or `rnorm()` functions. You can choose any length for these variables as per your choice. You can refer to the following [`modelr` page](https://modelr.tidyverse.org/reference/index.html) for reference. 

1. Fit a linear model `(y ~ x)` using the `lm()` function to predict `y` using `x`.

2. Using the `modelr` function `add_predictions()`, generate prediction values for your model.

3. Using the `modelr` function `add_residuals()`, generate residual values for your model.

4. Using the appropriate `ggplot` functions plot your predictions.

5. Using the appropriate `ggplot` functions plot the residuals for your model.

### Generate Data

```{r}
set.seed(496)
df <- tibble(`x` = sort(runif(15)), `y` = runif(15))

df
```

#### Assignment 1

1. Fit a linear model (y~x) using the `lm()` function to predict `y` using `x`.

```{r}
fit <- lm(y ~ x, data = df)
```

#### Assignment 2

2. Using the `modelr` function `add_predictions()`, generate prediction values for your model.

```{r}
df <- df %>% 
  add_predictions(model = fit, var = "pred")

df
```

#### Assignment 3

3. Using the `modelr` function `add_residuals()`, generate residual values for your model.

```{r}
df <- df %>% 
  add_residuals(model = fit, var = "resid")

df
```

#### Assignment 4

4. Using the appropriate `ggplot` functions plot your predictions.

```{r}
ggplot(data = df, aes(x = x, y = pred)) +
  geom_point() +
  geom_point(aes(y = y), alpha = 0.2, color = "red") +
  labs(title = "Plot of Independent Variable and Predicted Values", subtitle = "Actual Y Values Plotted in Red", x = "Independent Variable (sort(runif(15))", y = "Predicted Values by add_predictions()", caption = "Data Generated with set.seed(496)")
```

#### Assignment 5

5. Using the appropriate `ggplot` functions plot the residuals for your model.

```{r}
ggplot(data = df, mapping = aes(x = x, y = resid)) +
  geom_hline(yintercept = mean(df$resid), color = "red", alpha = 0.5) +
  geom_point() +
  labs(title = "Plot of Independent Variable and Residuals", subtitle = "Mean Residual Value Plotted in Red", x = "Independent Variable (sort(runif(15))", y = "Residual Values by add_residuals()", caption = "Data Generated with set.seed(496)")
```
