---
title: "Week 6 Lab Assignment 2"
output: 
  html_document:
    df_print: paged
author: "Jonathan Bonaguro"
date: "February 28, 2020"
---

```{r, echo = FALSE, message = FALSE}
library(tidyverse)
```

## Assignments

You can refer to the following refernce page for answering these questions. <https://readr.tidyverse.org/reference/index.html>

1. Using appropriate readr family of function read a dataset(from either a .txt or .csv file of your choice) and store it in a tibble. Attach the .txt or CSV file during your submission.
2. Now using the appropriate readr family of function write a tibble object of your choice to a CSV file. Attach the CSV file you created for submission.
3. Create an appropriate vector and use the following functions to successfully parse that vector.
    a. parse_logical
    b. parse_integer
    c. parse_double
    d. parse_vector
4. Create a vector of
    a. numeric and
    b. logical type
    c. and use the guess_parser() function to return output.
5. Create a vector of
    a. numeric and
    b. logical type
    c. and use the parse_guess() function to return the output.

You can refer to the following reference page for answering these questions. <https://tidyr.tidyverse.org/reference/pivot_wider.html>

6. Load the relign_income dataset. The pivot_longer() function is an updated version of gather() function you got used to in the hands-on Lab. Use the pivot_longer function to tidy this dataset.
7. Use the appropriate tidyr function to organize the untidy language_diversity data (attached csv file) into a tidy version.

#### Assignment 1

1. Using appropriate readr family of function read a dataset(from either a .txt or .csv file of your choice) and store it in a tibble. Attach the .txt or CSV file during your submission.

```{r}
russ_demo <- read_csv("Data/russian_demography_data.csv", col_types = cols(region = col_character()))
russ_demo
```

#### Assignment 2

2. Now using the appropriate readr family of function write a tibble object of your choice to a CSV file. Attach the CSV file you created for submission.

```{r}
# create dataset of demography data for Nizhny Novgorod Oblast
nizhny_novgorod_data <- russ_demo %>% 
  filter(region == "Nizhny Novgorod Oblast")

write_csv(nizhny_novgorod_data, "Data/nizhny_novgorod_demo_data.csv")
```

#### Assignment 3

3. Create an appropriate vector and use the following functions to successfully parse that vector.
    a. parse_logical
    b. parse_integer
    c. parse_double
    d. parse_vector

```{r}
vec <- rep(as.character(c(1.0, 0.0, 1.0, 1.0, 0.0, 0.0, 1.0, 0.0)), times = 10)

parse_logical(vec)
parse_integer(vec)
parse_double(vec)
parse_vector(vec, collector = col_logical())
```

#### Assignment 4

4. Create a vector of
    a. numeric and
    b. logical type
    c. and use the guess_parser() function to return output.

```{r}
num_vector <- c(1:5, 7:20, 8:1)
logical_vector <- c(TRUE, FALSE, TRUE, TRUE, FALSE, FALSE, TRUE, FALSE)

guess_parser(num_vector)
guess_parser(logical_vector)
```

#### Assignment 5

5. Create a vector of
    a. numeric and
    b. logical type
    c. and use the parse_guess() function to return the output.

```{r, error=TRUE}
num_vector <- as.character(c(1:5, 7:20, 8:1))
logical_vector <- as.character(c(TRUE, FALSE, TRUE, TRUE, FALSE, FALSE, TRUE, FALSE))

parse_guess(num_vector)
parse_guess(logical_vector)
```

#### Assignment 6

6. Load the relign_income dataset. The pivot_longer() function is an updated version of gather() function you got used to in the hands-on Lab. Use the pivot_longer function to tidy this dataset.

```{r}
relig_income

relig_income %>% 
  pivot_longer(-religion, names_to = "income", values_to = "count")
```

#### Assignment 7

7. Use the appropriate tidyr function to organize the untidy language_diversity data (attached csv file) into a tidy version.

```{r, message = FALSE}
language_diversity <- read_csv("Data/language_diversity.csv")

language_diversity

language_diversity %>% 
  pivot_wider(names_from = Measurement, values_from = Value) %>% 
  arrange(Continent)

```

