---
title: "Week 6 Lab Assignment 1"
output: 
  html_document:
    df_print: paged
author: "Jonathan Bonaguro"
date: "February 26, 2020"
---

```{r, echo = FALSE, message = FALSE}
library(tidyverse)
```

## Assignments

1. Using appropriate base R’s read family of functions read a dataset (from either a .txt or .csv file of your choice) and store it in a dataframe. Attach the txt or CSV file during your submission.
2. Now using the appropriate base R’s write family of functions write a dataframe object to a CSV file of your choice. Attach the CSV file you created for submission.

**Use the tibble reference page (https://tibble.tidyverse.org/reference/index.html) for help for the following questions:**

3. Create a tibble of your choice.
    a. Now add a row of new data to this tibble.
    b. Now add a column of new data to this tibble.
4. Create a tribble of your choice.
5. Create a R object with vectors of single type of your choice.
    a. Now, convert this R object to a tibble.
    b. Convert the tibble back to the original vector type.
6. Create a list object and convert it to a tibble.
7. Convert the mtcars dataframe to a tibble.
8. Using the pipe operator transform the mtcars to a tibble and check
    a. whether the mtcars tibble has rownames?
    b. whether the mtcars tibble has colnames?
9. Convert the rownames of the mtcars to a separate column so it can be retained within the tibble format and store this new object as a tibble

### Assignment 1 

1. Using appropriate base R’s read family of functions read a dataset (from either a .txt or .csv file of your choice) and store it in a dataframe. Attach the txt or CSV file during your submission.

```{r}
deck <- read.csv("deck.csv", header = TRUE)
deck
```

### Assignment 2

2. Now using the appropriate base R’s write family of functions write a dataframe object to a CSV file of your choice. Attach the CSV file you created for submission.

```{r}
spades <- deck[deck$suit == "spades",]
write.csv(spades, file = "spades.csv", row.names = FALSE)
```

### Assignment 3

3. Create a tibble of your choice.
    a. Now add a row of new data to this tibble.
    b. Now add a column of new data to this tibble.

```{r}
my_courses <- tibble(subject = c("PS", "PS", "STAT", "PSYCH"), number = c(328, 396, 385, 496))

my_courses <- my_courses %>% 
  add_row(subject = "PS", number = 496)

my_courses <- my_courses %>% 
  add_column(instructor = c("Ksiazkiewicz", "Prorok", "Ngyuen", "Adhimoolam", "Althuas"))

my_courses
```

### Assignment 4

4. Create a tribble of your choice.

```{r}
last_semester_courses <- tribble(
  ~subject, ~number, ~instructor,
  "STAT", 400, "Chatterjee",
  "STAT", 107, "Fagen-Ulmschneider",
  "SOC", 280, "Akresh",
  "PS", 495, "Winters"
)

last_semester_courses
```

### Assignment 5

5. Create a R object with vectors of single type of your choice.
    a. Now, convert this R object to a tibble.
    b. Convert the tibble back to the original vector type.

```{r}
random_vector <- c(a = 1, b = 2, c = 3)

random_vector <- enframe(random_vector)

random_vector <- deframe(random_vector)

random_vector
```

### Assignment 6

6. Create a list object and convert it to a tibble.

```{r}
my_list <- list(first = c("Jonathan", "Kae"), middle = c("Carlyle", "Nicole"), last = "Bonaguro")
my_list
as_tibble(my_list, .name_repair = "minimal")
```

### Assignment 7

7. Convert the mtcars dataframe to a tibble.

```{r}
as_tibble(mtcars)
```

### Assignment 8

8. Using the pipe operator transform the mtcars to a tibble and check
    a. whether the mtcars tibble has rownames?
    b. whether the mtcars tibble has colnames?

```{r}
mtcars %>% 
  as_tibble() %>% 
  has_rownames()

mtcars %>% 
  as_tibble() %>% 
  colnames()
```

### Assignment 9

9. Convert the rownames of the mtcars to a separate column so it can be retained within the tibble format and store this new object as a tibble

```{r}
mtcars_tibble <- mtcars %>% 
  as_tibble() %>% 
  rownames_to_column(var = "rownames")

mtcars_tibble
```

