---
title: "Week 8 Lab Assignment 2"
output: 
  html_document:
    df_print: paged
author: "Jonathan Bonaguro"
date: "March 13, 2020"
---

```{r setup environment, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(collapse = TRUE)
library(nycflights13)
library(tidyverse)
library(htmlwidgets)
```

## Assignments

Use only regular expressions for pattern matching to answer the questions below. You can refer to the reference page on regular expressions [here.](https://stringr.tidyverse.org/articles/regular-expressions.html) 

1. Using the words dataset:
    a. Return all the words beginning with the letter “s” or “a” and ending with any character.
    b. Return all the words that does not start with a vowel.
    c. Find the proportion of words that have repeat of any character (eg.,cc in accept).
    d. Find all the words containing exactly two vowels and count them.
    e. Find all the words that starts with the same character and ends with the same character.

Combine your knowledge on factors and datetime functions to answer the following questions:

a. Using the flight data set create a date time variable for departure time from the year, month, day, and dep_time variables.

b. From the date time object create a weekday variable, and calculate the total number of flights departing on each day of the week througout the year.

c. Now create factors by treating each day of the week as a level and calculate the mean departure delay across each day of the week throughout the year.

#### Assignment 1

1. Using the words dataset:
    a. Return all the words beginning with the letter “s” or “a” and ending with any character.
    b. Return all the words that does not start with a vowel.
    c. Find the proportion of words that have repeat of any character (eg.,cc in accept).
    d. Find all the words containing exactly two vowels and count them.
    e. Find all the words that starts with the same character and ends with the same character.

```{r assignment 1}
# a. Return all the words beginning with the letter “s” or “a” and ending with any character.
str_subset(words, pattern = "^s|^a")

# b. Return all the words that does not start with a vowel.
str_subset(words, pattern = "^[^aeioy]")

# c. Find the proportion of words that have repeat of any character (eg.,cc in accept).
length(str_subset(words, pattern = "(.)\\1")) / length(words)

# d. Find all the words containing exactly two vowels and count them.
sum(str_count(words, pattern = "[aeioy]") == 2)

# e. Find all the words that starts with the same character and ends with the same character.
str_subset(words, pattern = "^(.).*\\1$")

```

#### Assignment 2

a. Using the flight data set create a date time variable for departure time from the year, month, day, and dep_time variables.

b. From the date time object create a weekday variable, and calculate the total number of flights departing on each day of the week througout the year.

c. Now create factors by treating each day of the week as a level and calculate the mean departure delay across each day of the week throughout the year.

```{r assignment 2}
flights <- flights %>% 
  mutate(dep_date_time = make_date(year, month, day))

flights %>% 
  mutate(weekday = wday(dep_date_time)) %>% 
  group_by(weekday) %>% 
  summarise(num_flights = n())

flights %>% 
  mutate(weekday = factor(wday(dep_date_time))) %>% 
  group_by(weekday) %>% 
  summarise(mean_dep_delay = mean(dep_delay, na.rm = TRUE))

```


