---
title: "Homework Assignment One"
author: "Jonathan Bonaguro"
date: "February 23, 2020"
output: 
  html_notebook:
    df_print: paged
---

## For the following homework, you can use the lecture slides, hands-on-notebooks, reference books, google search, and `?` in R console as resources for help.

### Problem 1

1. (1 Point) What type of vector values are stored in object `a`. `a <- c("1","2","3")`. Go ahead and convert them to `numeric` type.

```{r}
a <- c("1","2","3")
typeof(a)

as.numeric(a)
```

### Problem 2

2. (1 Point) In R, the following arthimetic operation `b*2` on the object `b<-c(4,5,5,7)` will result in the following solution `(8,10,10,14)`. Which of the following accurately descibes how R handled this operation: 
    a) Arithmetic recyling
    b) **Vector recyling**
    c) Complex substituion
    d) Vector coercion
    
### Problem 3

3.  (2 Points) Let’s say you want to store multiple vector types in a R object `c`. What is wrong with the following code: `c <- c("Good Morning",1,TRUE)`. Can you rewrite this code appropriately?

**`c` automatically concatenates a vector, and vectors have to be data of the same type. The following code records the data as a type `list`**

```{r}
c <- c(list("Good Morning", 1, TRUE))
c
```

### Problem 4

4. (1 Point) Which of the following describes accurately about `class` of a R object?
    a) type
    b) dimension
    c) **Attribute**
    d) None of the above
    
### Problem 5

5. (1 Point) Which among the following operation will return the entire `mtcars` dataframe (including all rows and colums):
    a) **`mtcars[TRUE,]`**
    b) `mtcars[TRUE,c(FALSE)]`
    c) `mtcars[TRUE,FALSE]`
    d) `mtcars[FALSE,TRUE]`

```{r}
mtcars[TRUE, ]
```

### Problem 6

```{r}
list_a <- list(1:2, "Harry Potter", c(TRUE, TRUE, TRUE))

list_a
```


6. In `list_a`, which of the following operation will return the value 3?
    a) `sum(list_a[[1]])`
    b) `sum(list_a[[3]])`
    c) `sum(list_a[1])`
    d) **both a and b are true**

### Problem 7

7. (5 points) What will the following combinations of logical and boolean operations return?
    a. `(2<1) & (3<4) & (5<6)` will return `FALSE`
    b. `(2<1) | (3<4) | (5<6)` will return `TRUE`
    c. `(FALSE<TRUE) & (TRUE>FALSE) & (FALSE!=TRUE)` will return `TRUE`
    d. `any(c(2<1 & 3<4 & 5<6),(2<1 | 3<4 | 5<6),(FALSE<TRUE & TRUE>FALSE & FALSE!=TRUE))` will return `TRUE`
    e. `all(1%in%c(2,3,4),0%in%c(FALSE,TRUE),1%in%100==100)` will return `FALSE`

### Problem 8

8. (2 Points) What is the use of `ifelse()` function? Execute the `ifelse()` function with an example of your choice.

**`ifelse()` exectues a logical test and returns the value of argument `yes=` if `test` returns `TRUE` and returns the value of argument `no=` if `test` returns `FALSE`.**

```{r}
x <- 1:10
sum(ifelse(x>=0, yes = x, NA))

x <- -2:10
sum(ifelse(x>=0, yes = x, NA))
```

### Problem 9

9. (1 Point) What is the use of `switch()` function?

**The `switch()` function basically takes the place of nested `else if` loops. It allows us to choose an argument to use based on an input, `EXPR`, which is then matched to defined arguments to determine what should be done.**

### Problem 10

10. (2 Points) What is the role of next and break statements in controling the flow of code? Write an example code of your choice to demonstrate the use of both next and break statements.

**`next` controls the flow by stopping the current iteration and advances the index of the loop. `break` exits a loop and transfer is moved to the next statement outside of the loop.**

```{r}
for(i in 1:10) {  
  if(i == 3) {
    next
  }
  if(i == 9) {
    break
  }
  print(paste("This is the number", i))
}

```

### Problem 11

11. (2 Points) What is the use of `sample()` function? Execute the `sample()` function with an example of your choice.

**`sample()` is used to take a sample of a size `size=` from an element `x=` (usually a vector)**

```{r}
set.seed(385)
sample(x = 1:100, size = 50)
```

### Problem 12

12. (2 Points) What is the use of `rep()` function? Execute the `rep()` function with an example of your choice.

**The `rep()` function replicates the values in an input vector `x =`. The output can be specified by either `times =` (the number of times to repeat `x`), or `length.out =` (the desired length of the output).**

```{r}
y <- 1:10
rep(x = y, length.out = 52)
rep(x = y, times = 3)
```

### Problem 13

13. (2 Points) What is the use of `runif()` function? Execute the `runif()` function with an example of your choice.

**`runif()` generates `n` random numbers on a uniform distribution from an interval specified by the arguments `min =` and `max =`. To demonstrate uniformity, I plot a histogram of the result of a `runif()` function on the interal `1:10`.**

```{r}
set.seed(385)
random <- runif(1000000, min = 0, max = 10)
hist(random, main = "Histogram of Uniform Distribution Generated by `runif()`", probability = TRUE, xlab = "Random Numbers Generated by `runif()`", breaks = 10)
```

### Problem 14

14. (2 Points) What is the use of `rnorm()` function? Execute the `rnorm()` function with an example of your choice.

**`rnorm` generates `n` random numbers from a normal distribution specified by `mean =` and `sd =`. To demonstrate normality, I plot a histogram of a `rnorm()` sample from the standard normal distribution.**

```{r}
set.seed(385)
random <- rnorm(1000000, mean = 0, sd = 1)
hist(random, main = "Histogram of Standard Normal Distribution Generated by `rnorm()`", probability = TRUE, xlab = "Random Numbers Generated by `rnorm()`", xlim = c(-3, 3))
```
