---
title: "Homework Assignment One"
author: "Jonathan Bonaguro"
date: "February 23, 2020"
output: 
  html_notebook:
    df_print: paged
---

## For the following homework, you can use the lecture slides, hands-on-notebooks, reference books, google search, and `?` in R console as resources for help.

### Problem 1

1. (1 Point) What type of vector values are stored in object `a`. `a <- c("1","2","3")`. Go ahead and convert them to `numeric` type.

```{r}
a <- c("1","2","3")
typeof(a)

as.numeric(a)
```

### Problem 2

2. (1 Point) In R, the following arthimetic operation `b*2` on the object `b::b<-c(4,5,5,7)` will result in the following solution `(8,10,10,14)`. Which of the following accurately descibes how R handled this operation: 
    a) Arithmetic recyling
    b) **Vector recyling**
    c) Complex substituion
    d) Vector coercion
    
### Problem 3

3.  (2 Points) Let’s say you want to store multiple vector types in a R object `c`. What is wrong with the following code: `c <- c("Good Morning",1,TRUE)`. Can you rewrite this code appropriately?

**`c` automatically concatenates a vector, and vectors have to be data of the same type. The following code records the data as a type `list`**

```{r}
c <- c(list("Good Morning", 1, TRUE))
c
```

### Problem 4

4. (1 Point) Which of the following describes accurately about `class` of a R object?
    a) type
    b) dimension
    c) **Attribute**
    d) None of the above
    
### Problem 5

5. (1 Point) Which among the following operation will return the entire `mtcars` dataframe (including all rows and colums):
    a) **`mtcars[TRUE,]`**
    b) `mtcars[TRUE,c(FALSE)]`
    c) `mtcars[TRUE,FALSE]`
    d) `mtcars[FALSE,TRUE]`

```{r}
mtcars[TRUE, ]
```

### Problem 6

6. In the following list `list_a`, which of the following operation will return the value 3?
    a) `sum(list_a[[1]])`
    b) `sum(list_a[[3]])`
    c) `sum(list_a[1])`
    d) **both a and b are true**

### Problem 7

7. (5 points) What will the following combinations of logical and boolean operations return?
    a. `(2<1) & (3<4) & (5<6)` == `FALSE`
    b. `(2<1) | (3<4) | (5<6)` == `TRUE`
    c. `(FALSE<TRUE) & (TRUE>FALSE) & (FALSE!=TRUE)` == `TRUE`
    d. `any(c(2<1 & 3<4 & 5<6),(2<1 | 3<4 | 5<6),(FALSE<TRUE & TRUE>FALSE & FALSE!=TRUE))` == `TRUE`
    e. `all(1%in%c(2,3,4),0%in%c(FALSE,TRUE),1%in%100==100)` == `FALSE`

### Problem 8

8. (2 Points) What is the use of `ifelse()` function? Execute the `ifelse()` function with an example of your choice.

**`ifelse()` exectues a logical test and returns the value of argument `yes=` if `test` returns `TRUE` and returns the value of argument `no=` if `test` returns `FALSE`.**

```{r}
x <- 1:10
sum(ifelse(x>=0, yes = x, NA))

x <- -2:10
sum(ifelse(x>=0, yes = x, NA))
```

### Problem 9

9. (1 Point) What is the use of `switch()` function?

