---
title: "Week 5 Lab Assignment 2"
output: 
  html_document:
    df_print: paged
author: "Jonathan Bonaguro"
date: "February 21, 2020"
---

## Assignment 

Create a function with traditional if else/for loop of your choice. You can then proceed to vectorize this code and rewrite this function in a vectorized manner. Use the `system.time()` function on both these functions to demonstrate that your code performs faster with the vectorization method.

### For Loop 

* Defined a `for` loop to print a vector of card colors based on a vector of card suits

```{r}
change_symbols <- function(vec){
  for (i in 1:length(vec)){
    if (vec[i] %in% c("hearts", "diamonds")) {
      vec[i] <- "red"
    } else if (vec[i] %in% c("spades", "clubs")) {
      vec[i] <- "black"
    } else {
      vec[i] <- NA
    } 
  }
  vec
}

draw <- c("hearts", "spades", "diamonds", "clubs", "joker")

many <- rep(draw, 1000000)

change_symbols(draw)

system.time(change_symbols(many))


```

### Vectorized Function

* Vectorized function to print a vector of card colors based on their suits

```{r}
change_vec <- function (vec) {
  vec[vec %in% c("hearts", "diamonds")] <- "red"
  vec[vec %in% c("spades", "clubs")] <- "black"
  vec[!vec %in% c("red", "black")] <- NA
  vec
}

change_vec(draw)

system.time(change_vec(many))
```

### Vectorized Function with Lookup Table

* Vectorized function using a lookup table to print a vector of card colors based on their suits

```{r}
change_vec2 <- function(vec){
  tb <- c("hearts" = "red", "diamonds" = "red", "spades" = "black", "clubs" = "black")
  unname(tb[vec])
}

change_vec2(draw)

system.time(change_vec2(many))
```

