---
title: "Week 4 Lab Assignment 1"
output: prettydoc::html_pretty
author: "Jonathan Bonaguro"
date: "February 12, 2020"
---

## Assignments 

1. Write an if-else statement that will take a numeric value and evaluate it (if it is less than, greater than, equal to zero) and print this information into the console.

2. Write an if-else statement to check if a number is an odd number or even number and print the information to the console. [Hint: use the modulus operator in R for this assignment].

3. Write a for loop that takes any R object containing vector of numeric values and returns it with every element of the values multipled by 2.

4. Write a for loop that takes the mtcars dataframe and prints the type of vector values contained in each column.

5. Create a matrix containing numeric values in 10 rows and 10 columns. Now write a for loop that will return the sum for each of the columns in this matrix.

6. Combine the if-else statement and for loop for this exercise . You can use the mtcars dataset for this assignment. Across the columns of this dataset check if they have a missing value and print “found missing values” if any of the columns have missing value. If no missing values are found print “no missing values”.

#### Assignment 1

1. Write an if-else statement that will take a numeric value and evaluate it (if it is less than, greater than, equal to zero) and print this information into the console.

```{r}
x <- 3
if(x > 5) {
  print("x is greater than 5")
} else if (x < 5) {
  print ("x is less than 5")
} else {
  print ("x is equal to 5")
}
```

#### Assignment 2

2. Write an if-else statement to check if a number is an odd number or even number and print the information to the console. [Hint: use the modulus operator in R for this assignment].

```{r}
x <- 5203
if((x %% 2) == 0) {
  print("x is an even number")
} else {
  print("x is an odd number")
}
```

#### Assignment 3

3. Write a for loop that takes any R object containing vector of numeric values and returns it with every element of the values multipled by 2.

```{r}
vector <- seq(from = 2, to = 10, by = 2)
output <- NA
for (i in 1:length(vector)) {
  output[i] <- (vector[i] *2)
}
output
```

#### Assignment 4

4. Write a for loop that takes the mtcars dataframe and prints the type of vector values contained in each column.

```{r}
output <- NA
for (i in 1:ncol(mtcars)) {
  output[i] <- class(mtcars[ , i])
}
output
```

#### Assignment 5

5. Create a matrix containing numeric values in 10 rows and 10 columns. Now write a for loop that will return the sum for each of the columns in this matrix.

```{r}
matrix <- matrix(1:100, nrow = 10, byrow = T)
output <- NA
for (i in 1:ncol(matrix)) {
  output[i] <- sum(matrix[ , i])
}
output
```

#### Assignment 6

6. Combine the if-else statement and for loop for this exercise . You can use the mtcars dataset for this assignment. Across the columns of this dataset check if they have a missing value and print “found missing values” if any of the columns have missing value. If no missing values are found print “no missing values”.

```{r}
mtcars[4, 5] <- NA
output <- matrix(NA, nrow = nrow(mtcars), ncol = ncol(mtcars), byrow = FALSE)
columns_missing_data <- NA
for (i in 1:ncol(mtcars)) {
  output[ , i] <- is.na(mtcars[ , i])
  for (j in 1:ncol(mtcars)){
    if (TRUE %in% output[ , j]) {
      columns_missing_data[j] <- "found missing values"
    } else {
       columns_missing_data[j] <- "found no missing values"
    }
  }
}
columns_missing_data
```
