---
title: "Week 4 Lab Assignment 2"
output: prettydoc::html_pretty
author: "Jonathan Bonaguro"
date: "February 14, 2020"
---

## Assignments

1. Write your own function that uses an explicit return() statement in the body of the code.
2. Write a function that uses stopifnot() in the body of code. Use the arguments in such a way that this function should return a error message if the specific conditions are not met within the arguments or body of the code.
3. Use the dot-dot-dot(…) argument in a function to return the output as list whenever you input a vector of numeric type.
4. Using the generic print() function, create your own custom S3 method that takes a R object of your choice and prints the output as per your choice.

#### Assignment 1

1. Write your own function that uses an explicit return() statement in the body of the code.

```{r}
sd_of_vector <- function(vect) {
  s <- sd(vect, na.rm = TRUE)
  return(s)
  print(paste("The standard deviation of the vector is", s))
}

sd_of_vector(c(1, 3, 5, NA, 12, 4))
```

#### Assignment 2

2. Write a function that uses stopifnot() in the body of code. Use the arguments in such a way that this function should return a error message if the specific conditions are not met within the arguments or body of the code.

```{r, error = TRUE}
sd_of_vector <- function(vect) {
  stopifnot(is.numeric(vect), all(is.na(vect) == FALSE))
  s <- sd(vect)
  return(s)
}

sd_of_vector(c(1, 3, 5, NA, 12, 4))
```

#### Assignment 3

3. Use the dot-dot-dot(…) argument in a function to return the output as list whenever you input a vector of numeric type.


```{r}
return_list <- function(...) {
  list(...)
}

return_list(c(1, 3, 5, NA, 12, 4), 1:10, 1:2)
```

#### Assignment 4

4. Using the generic print() function, create your own custom S3 method that takes a R object of your choice and prints the output as per your choice.

```{r}
a <- as.list(c(1, 3, 5, NA, 12, 4))
class(a) <- c("custom_print")

print.custom_print<-function(x){
  print(paste("Your list contains",x))
}

print(a)
b <- list(1, 10)
class(b) <- c("custom_print")
b
```

