---
title: "STAT385 Notes Week 3"
author: "Jonathan Bonaguro"
output:
  html_document:
    df_print: paged
date: "Week 3"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Monday, Feb. 3, 2020

#### Changing values in place

To modify a value within an R object:

* First, describe the value you wish to modify
* Use the assignment operator `<-` to overwrite those values
* R will update the selected values **i the original object**

```{r}
x <- c(0,0,0,0,0)
x[1] <- 100
x
```

To change multiple values at once, make sure the number of **new values** equals the number of **selected values**:

```{r}
x[c(1,3,5)] <- c(5,5,5)
x
x[4:6] <- x[4:6] + 1
x
```

Just like populating vectors, R will repeat itself until it can't anymore.

```{r, error=TRUE}
x[1:2] <- c(2)
x
x[1:2] <- c(100,100,100)
x
```

#### Looking at Data Frames

Let's add a new column to `deck`:

```{r}
deck <- read.csv("https://nkha149.github.io/stat385-sp2020/files/data/cards.csv")
deck$new_column <- 1:52
deck
```

Lets change some values:

```{r}
deck[c(13, 26, 39, 52), 3] <- 14
deck
```

#### Logical Operators

```{r}
1 > 2
1 > c(0:2)
c(1,2,3) == c (3,2,1)
1 %in% c(3:5)
```

The operators are standard. A different one, `%in%` looks if values are in a group. This operator does not follow element-wise execution. `%in%` independently tests whether each value is somewhere in the vector. 

```{r}
c(1:4) %in% c(3:5)
```

**Note:** `=` is an assignment operator like `<-`. Make sure to use `==`.

#### Using Logical Operator for Subsetting

* Let's say in a certain game, aces receive the highest value, say 14.
* Modify the values in `deck` to follow this rule
* What are the values of the aces in `deck` right now? 

```{r}
deck$face
deck$face == "ace"
```

Columns that say `TRUE` are aces. Let's subet. 

```{r}
deck[deck$face == "ace",]
```

Now, let's update the value for the aces to be 14!

```{r}
deck[deck$face == "ace", c("value")] <- 14
deck[deck$face == "ace",]
```

#### Boolean Operators 

* And `&` 
* Or `|`
* Exclusive or `xor`
* Not `!`
* Any `any`
* All `all`  

To use Boolean operator, place it between two *complete* logical tests.  

```{r}
a <- c(1:3)
b <- c(1:3)
c <- c(1,2,4)
a == b
c == b
a == c
```

#### Excercise 1 

Read in data from csv:

```{r}
hearts_deck <- read.csv("https://nkha149.github.io/stat385-sp2020/files/data/cards.csv")
```

In a new game called **Hearts**, every card has the value `0` except for the suit Hearts and the Queen of spades. Set up the deck in this way:

* Hearts all have value `1`
* The Queen of Spades has value `13`
* Save data as Data Frame `hearts_deck`

```{r}
# Set all values in deck equal to 0
hearts_deck$value <- 0
hearts_deck[hearts_deck$suit == "hearts", 3] <- 1
hearts_deck[hearts_deck$suit == "spades" & hearts_deck$face == "queen", 3] <- 13
hearts_deck
```

Ta da!

## Friday, Feb. 7, 2020

#### `if` Statement

* an `if` statement tells R to do a certain task for a certain case
* `condition` object should be a logical test
* if `condition` is `TRUE`
    + R will run all the code that appears in the `{}` following the `if` statement
* if `condition` is `FALSE`
    + R will skip the curly brackets

```{r}
num <- -2

if (num < 0) {
  num <- num * -1
}

num
```

```{r}
num <- 4

if (num < 0) {
  num <- num * -1
}

num
```

#### `else` statement

* `else` tells R what to do when the condition is `FALSE`

```{r}
my_round <- function(a){
  decimal <- a - trunc(a)
  if (decimal >= 0.5) {
    a <- trunc(a) + 1
  } else {
    a <- trunc(a)
  }
  return(a)
}
my_round(3.14)
```

#### `if else` ladder

```{r}
who_win <- function(a, b){
  if (a > b) {
    print("A wins!")
  } else if (a <b) {
    print("B wins!")
  } else {
    print("Tie.")
  }
}
who_win(3, 1)
who_win(5, 10)
who_win(3, 3)
```

#### Histograms

* Basic histogram using `hist()`
    + Change the plot title with `main =`
    + Change x-axis label with `xlab=`

```{r}
hist(x = cars$speed, main = "Histogram of Speed", xlab = "Speed (mph)")
```

* Change histogram from frequency to probability using `probability = TRUE`

```{r}
hist(x = cars$speed, main = "Histogram of Speed", xlab = "Speed (mph)", probability = TRUE)
```

* Change the break points in the histogram celss using `breaks = `
    + Use `seq()` to compute vector of breakpoints

```{r}
hist(x = cars$speed, main = "Histogram of Speed", xlab = "Speed (mph)", probability = TRUE, breaks = seq(from = 0, to = 32, by = 4))
```

* Provide `breaks =` a single number to specify breaks

```{r}
hist(x = cars$speed, main = "Histogram of Speed", xlab = "Speed (mph)", probability = TRUE, breaks = 10)
```

* Color inside of bins with `col =`

```{r}
hist(x = cars$speed, main = "Histogram of Speed", xlab = "Speed (mph)", probability = TRUE, breaks = seq(from = 0, to = 32, by = 4), col = "darkorange")
```

* Color border of histogram using `border =`

```{r}
hist(x = cars$speed, main = "Histogram of Speed", xlab = "Speed (mph)", probability = TRUE, breaks = seq(from = 0, to = 32, by = 4), border = "dodgerblue")
```

* Add box and grid using `box()` and `grid`

```{r}
hist(x = cars$speed, main = "Histogram of Speed", xlab = "Speed (mph)", probability = TRUE, breaks = seq(from = 0, to = 32, by = 4))
box()
grid()
```

#### Boxplot

* Used for exploratory data analysis 

```{r}
boxplot(x = cars$speed)
```

* Add title and axis label

```{r}
boxplot(x = cars$speed, main = "Boxplot of Speed", ylab = "Speed (mph)")
```

* Rotate plot using `horizontal = TRUE`

```{r}
boxplot(x = cars$speed, main = "Boxplot of Speed", ylab = "Speed (mph)", horizontal = TRUE)
```

#### Scatterplot

* Use `plot()` to make a Scatterplot

```{r}
plot(x = cars$speed, y = cars$dist)
```

* Add title and labels

```{r}
plot(x = cars$speed, y = cars$dist, main = "Car Speed vs. Stopping Distance", xlab = "Speed (mph)", ylab = "Stopping Distance (ft)")
```

* Add color and grid. Change the plotting symbol using `pch =`

```{r}
plot(x = cars$speed, y = cars$dist, main = "Car Speed vs. Stopping Distance", xlab = "Speed (mph)", ylab = "Stopping Distance (ft)", col = "dodgerblue", pch = 17)
grid()
```

* Let's try to change the scale on the axis using `xlim =` and `ylim =`

```{r}
plot(x = cars$speed, y = cars$dist, main = "Car Speed vs. Stopping Distance", xlab = "Speed (mph)", ylab = "Stopping Distance (ft)", col = "dodgerblue", pch = 19, xlim = c(0, 30), ylim = c(0, 150))
grid()
```

#### Play around with plotting

* Plot histogram of `eruptions` in dataset `faithful`

```{r}
hist(faithful$eruptions, main = "Histogram of Old Faithful Eruption Length (min)", xlab = "Length of Time (min)", breaks = 20, probability = TRUE, border = "dodgerblue")
```

* Plot a scatterplot between the eruption time and waiting time to the next eruption

```{r}
plot(x = faithful$eruptions, y = faithful$waiting, main = "Old Faithful Eruption Time vs. Wait Time (min)", xlab = "Eruption Time (min)", ylab = "Waiting Time (min)", col = "dodgerblue", pch = 19)
grid()
```

* Another way to plot is with `formula =` and `data =`

```{r}
plot(formula = waiting ~ eruptions, data = faithful, main = "Old Faithful Eruption Time vs. Wait Time (min)", xlab = "Eruption Time (min)", ylab = "Waiting Time (min)", col = "dodgerblue", pch = 19)
grid()
```

* This is a preferred method for plotting, especially when we get to regressions. This specifies that `waiting` is being plotted as a function of `eruption`