---
title: "STAT385 Notes Week 3"
author: "Jonathan Bonaguro"
output:
  html_document:
    df_print: paged
date: "Week 3"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Monday, Feb. 3, 2020

#### Changing values in place

To modify a value within an R object:

* First, describe the value you wish to modify
* Use the assignment operator `<-` to overwrite those values
* R will update the selected values **i the original object**

```{r}
x <- c(0,0,0,0,0)
x[1] <- 100
x
```

To change multiple values at once, make sure the number of **new values** equals the number of **selected values**:

```{r}
x[c(1,3,5)] <- c(5,5,5)
x
x[4:6] <- x[4:6] + 1
x
```

Just like populating vectors, R will repeat itself until it can't anymore.

```{r, error=TRUE}
x[1:2] <- c(2)
x
x[1:2] <- c(100,100,100)
x
```

#### Looking at Data Frames

Let's add a new column to `deck`:

```{r}
deck <- read.csv("https://nkha149.github.io/stat385-sp2020/files/data/cards.csv")
deck$new_column <- 1:52
deck
```

Lets change some values:

```{r}
deck[c(13, 26, 39, 52), 3] <- 14
deck
```

#### Logical Operators

```{r}
1 > 2
1 > c(0:2)
c(1,2,3) == c (3,2,1)
1 %in% c(3:5)
```

The operators are standard. A different one, `%in%` looks if values are in a group. This operator does not follow element-wise execution. `%in%` independently tests whether each value is somewhere in the vector. 

```{r}
c(1:4) %in% c(3:5)
```

**Note:** `=` is an assignment operator like `<-`. Make sure to use `==`.

#### Using Logical Operator for Subsetting

* Let's say in a certain game, aces receive the highest value, say 14.
* Modify the values in `deck` to follow this rule
* What are the values of the aces in `deck` right now? 

```{r}
deck$face
deck$face == "ace"
```

Columns that say `TRUE` are aces. Let's subet. 

```{r}
deck[deck$face == "ace",]
```

Now, let's update the value for the aces to be 14!

```{r}
deck[deck$face == "ace", c("value")] <- 14
deck[deck$face == "ace",]
```

#### Boolean Operators 

* And `&` 
* Or `|`
* Exclusive or `xor`
* Not `!`
* Any `any`
* All `all`  

To use Boolean operator, place it between two *complete* logical tests.  

```{r}
a <- c(1:3)
b <- c(1:3)
c <- c(1,2,4)
a == b
c == b
a == c
```

#### Excercise 1 

Read in data from csv:

```{r}
cards <- read.csv("https://nkha149.github.io/stat385-sp2020/files/data/cards.csv")
```

In a new game called **Hearts**, every card has the value `0` except for the suit Hearts and the Queen of spades. Set up the deck in this way:

* Hearts all have value `1`
* The Queen of Spades has value `13`
* Save data as Data Frame `hearts_deck`

