---
title: "HS2_IS457_33"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Do not remove any of the comments. These are marked by #
# HW 2 - Due Monday, Sep 24 2018 in moodle and for non-online students hardcopy in class.
# (1). Please upload R code and report to Moodle with filename: HW2_IS457_YourCourseID.
# (2). Turn in hard copy of your report in class.

### Class ID: 33

# In this assignment you will practice manipulating vectors and dataframes.
# You will take subsets, create new data structures, and end with creating a fantastic plot.
# You will work with the CO2 dataset in the R library and a dataset called SFHousing. 
# Before beginning with the datasets however, you will do some warm up exercises.



# PART 1. Warm up (4 pts)
# Q1. Create a Vector like this (0 0 0 3 3 3 6 6 6 9 9 9 12 12 12 15 15 15 18 18 18) 
#     with functions seq() and rep() and call it "vec" (1 pt)

### Your code below
```{r}
vec=rep(seq(0, 18, by = 3), each = 3)
```

# Q2. Calculate the fraction of elements in vec that are more than or equal to 9. (2 pts)
# hint: R can do vectorized operations. 

### Your code below
```{r}
length(which(vec>=9))/length(which(vec>=0))
```

# Q3. Create a Vector like this (1 2 2 3 3 3 4 4 4 4 5 5 5 5 5)
#     with functions rep() and the : operator (1 pt)

### Your code below
```{r}
rep(1:5, 1:5)
```

# PART II. CO2 Data (9 pts)
# Q4. Use R to generate descriptions of the CO2 data which is already available with the base R installation (it
# is called CO2 in R. Please note that we are using the CO2 dataset and not the similarly named co2 dataset). 
# Print out the summary of each column and the dimensions of the dataset. (2 pts.)
# (hint: you may find the summary() and dim() useful). 
# Write up your descriptive findings and observations of the R output. (1 pt.)

### Your code below:
```{r}
dim(CO2)
summary(CO2)
help("CO2")
```

### Your answer below:
```{r}
# The CO2 datatset contains 84 ovservations of 5 variables. There are three factor variables (plant, type, Treatment)
# and two numeric variables (conc and uptake). The variable Plant creates a unique identifier for each plant, type
# gives us the origin of the plant, treatment tells us whether it was chilled or non chilled, conc gives us the ambient
# carbon dioxide concentrations, and uptake gives us the CO2 uptake rates. R gives us a summary of 42 plants from each
# region, and 42 treated both chilled and nonchilled. It also gives us min, max, median, mean, and 1st and 3rd
# quartile values of the numeric variables. 
```

# Q5. Show last 8 plants' uptake values (1 pt.)

### Your code below:
```{r}
CO2$uptake[77:84]
```

# Q6. Show all plants' uptake values except the first 20 plants'. (1 pt.)

### Your code below:
```{r}
CO2$uptake[-1:-20]
```

# Q7. Calculate the mean of uptake subseted by the "Treatment" variable.(1 pt)
# hint: apply function family.

### Your code below:
```{r}
tapply(CO2$uptake, CO2$Treatment, mean, na.rm=TRUE)
```

# Q8. Create a logical vector uptake_treatment . (2 pts)

# For the plants with Chilled treatment (Treatment == "chilled"), return value TRUE when uptake > 30.
# For the plants with Non-Chilled treatment (Treatment == "nonchilled"), return value TRUE when uptake > 40.

### Your code below:
```{r}
uptake_treatment=(CO2$Treatment=="chilled"&CO2$uptake>30)|(CO2$Treatment=="nonchilled"&CO2$uptake>40)
uptake_treatment
```  

# Q9. Here is an alternative way to create the same vector in Q8.
# First, we create a numeric vector uptake_test that is 30 for each plant with chilled treatment
# and 40 for each plant with non chilled treatement. To do this, first create a vector of length 2 called 
# test_val whose first element is 40 and second element is 30. (1 pt)

### Your code below:
```{r}
test_val=c(40,30)
```
# Create the uptake_test vector by subsetting test_val by position, where the 
# positions could be represented based on the Treatment column in CO2. (1 pt)

### Your code below
```{r}
uptake_test=test_val[CO2$Treatment]
```

# Finally, use uptake_test and the uptake column to create the desired vector, and
# call it uptake_treatment2. (1 pt)

### Your code below
```{r}
uptake_treatment2=CO2$uptake > uptake_test
uptake_treatment2
```
#PART 3.  San Francisco Housing Data (25 pts.)
#
# Load the data into R.
```{r}
load(url("https://www.stanford.edu/~vcs/StatData/SFHousing.rda"))
```
# Q10. (3 pts.)
# What objects are in SFHousing.rda? Give the name and class of each.

### Your code below
```{r}
class(cities)
class(housing)
```
### Your answer here
#There are two objects in SFHousing.rda- two data frames named cities and housing. 


# Q11. give a summary of each object, including a summary of each variable and the dimension of the object. (4 pts)

### Your code below
```{r}
summary(cities)
dim(cities)
summary(housing)
dim(housing)
```

# Q12. After exploring the data (maybe using the summary() function), describe in words the connection
# between the two objects (e.g., what links them together). (2 pts)

### Write your response here
```{r}
#That data have a few things in commonly, namely county of location as well as information on latitute and longitude.
```

# Q13. Describe in words two problems that you see with the data. (2 pts)

### Write your response here
```{r}
# The first problem I see with the data is that not all of the rows are equal length, which could make some analysis
# more difficult. The second problem I see is the high number of NA's recorded in the data set, which will 
# change analysis based on missing data. 
```

# Q14. (2 pts.)
# We will work with the houses in San Francisco, Fremont, Vallejo, Concord and Livermore only.

# Subset the housing data frame so that we have only houses in these cities
# and keep only the variables county, city, zip, price, br, bsqft, and year.

# Call this new data frame SelectArea. This data frame should have 36686 observations
# and 7 variables. (Note you may need to reformat any factor variables so that they
# do not contain incorrect levels)

### Your code below
```{r}
cities = c("San Francisco", "Fremont", "Vallejo", "Concord", "Livermore")
SelectArea = housing[housing$city %in% cities, c("county", "city", "zip", "price", "br", "bsqft", "year")]
dim(SelectArea)
```

# Q15. (3 pts.)
# We are interested in making plots of price and size of house, but before we do this
# we will further subset the housing dataframe to remove the unusually large values.

# Use the quantile function to determine the 95th percentile of price and bsqft
# and eliminate all of those houses that are above either of these 95th percentiles
# Call this new data frame SelectArea (replacing the old one) as well. It should 
# have 33693 observations.

### Your code below
```{r}
SelectArea=subset(SelectArea,(SelectArea$price<=quantile(SelectArea$price,c(.95), na.rm=TRUE)))
SelectArea=subset(SelectArea,(SelectArea$bsqft<=quantile(SelectArea$bsqft,c(.95), na.rm=TRUE)))
```

# Q16. (2 pts.)
# Create a new vector that is called price_per_sqft by dividing the sale price by the square footage
# Add this new variable to the data frame.

### Your code below
```{r}
SelectArea$price_per_sqft=SelectArea$price/SelectArea$bsqft
```

# Q17. (2 pts.)
# Create a vector called br_new, that is the number of bedrooms in the house, except
# when the number is greater than 5, set it (br_new) to 5.  

### Your code below
```{r}
SelectArea$br_new=SelectArea$br
SelectArea$br_new[SelectArea$br_new >= 5]=5
```

# Q18. (4 pts. 2 + 2 - see below)
# Use the heat.colors function to create a vector of 5 colors, call this vector rCols.
# When you call this function, set the alpha argument to 0.25.

# Create a vector called brCols where each element's value corresponds to the color in rCols 
# indexed by the number of bedrooms in the br_new.

# For example, if the element in br_new is 3 then the color will be the third color in rCols.
# (2 pts.)

### Your code below
```{r}
rCols=heat.colors(5, alpha=0.25)

SelectArea$brCols=NA
SelectArea$brCols[SelectArea$br_new == 1]=rCols[1]
SelectArea$brCols[SelectArea$br_new == 2]=rCols[2]
SelectArea$brCols[SelectArea$br_new == 3]=rCols[3]
SelectArea$brCols[SelectArea$br_new == 4]=rCols[4]
SelectArea$brCols[SelectArea$br_new == 5]=rCols[5]
```

######
# We are now ready to make a plot!
```{r}
plot(price_per_sqft ~ bsqft, data = SelectArea,
     main = "Housing prices in the San Francisco Area",
     xlab = "Size of house (square ft)",
     ylab = "Price per square foot",
     col = brCols, pch = 18, cex = 0.5)
legend(legend = 1:5, fill = rCols, "topright")
```

# what's your interpretation of the plot? 
# e.g., the trend? the cluster? the comparison? (1 pt.)
```{r}
#My interpretation of the plot is that the more bedrooms a house has, the bigger it typically is in terms of square feet.
#It also suggests that the less bedrooms it has, the more expensive the house is in terms of price per square foot.
```
