---
title: "Week 10 Lab Assignment 1"
output: 
  html_document:
    df_print: paged
author: "Jonathan Bonaguro"
date: "April 1, 2020"
---

```{r setup environment, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(collapse = TRUE)
library(tidyverse)
library(hexbin)
```

## Assignments

Use the diamonds dataset to answer the following questions. You can also refer to [`ggplot2` cheatsheet and reference page](https://ggplot2.tidyverse.org/reference/index.html).

1. Visualize the distribution of the price of the diamonds with the histogram (`geom_histogram`) and frequency polygon (`geom_freqpoly`). Make sure you try and explore variable binwidths for these visualizations.

2. Visualize the covariation of price of the diamonds in relation to the cut variable with a boxplot (`geom_boxplot`).
    a. Now reorder the above boxplot by using the median value of the price and visualize the boxplot.
    b. Now flip the coordinates of x and y and create a new boxplot.
    c. Visualize the covariation of the price of the diamonds in relation to both cut and clarity using boxplot.
    d. Overlay the boxplot with the original datapoints. Use the 2a question above for this plot.

3. Visualize the covariation of the price of the diamonds in relation to the cut variable by using the violin plot(`geom_violin`).
    a. Mark the following quartiles in the above violin plot (0.25,0.5,0.75).

4. Explore the covariation of the price and carat of diamonds using rectangular (`geom_bin2d`) and hexagonal bins (`geom_hex`).Make sure you explore this relationship with variable binwidths.

#### Assignment 1

1. Visualize the distribution of the price of the diamonds with the histogram (`geom_histogram`) and frequency polygon (`geom_freqpoly`). Make sure you try and explore variable binwidths for these visualizations.

```{r}
ggplot(data = diamonds, mapping = aes(x = price)) +
  geom_histogram(bins = 50) + 
  labs(title = "Histogram of Diamond Price", x = "Price (USD)", y = "Count", caption = "Diamonds data provided by ggplot2")

ggplot(data = diamonds, mapping = aes(x = price)) +
  geom_freqpoly(bins = 50) + 
  labs(title = "Frequency Diagram of Diamond Price", x = "Price (USD)", y = "Count", caption = "Diamonds data provided by ggplot2")

```

#### Assignment 2

2. Visualize the covariation of price of the diamonds in relation to the cut variable with a boxplot (`geom_boxplot`).
    a. Now reorder the above boxplot by using the median value of the price and visualize the boxplot.
    b. Now flip the coordinates of x and y and create a new boxplot.
    c. Visualize the covariation of the price of the diamonds in relation to both cut and clarity using boxplot.
    d. Overlay the boxplot with the original datapoints. Use the 2a question above for this plot.

```{r}
ggplot(data = diamonds, mapping = aes(x = cut, y = price)) +
  geom_boxplot(outlier.alpha = 0.1) + 
  labs(title = "Box Plot of Diamond Price Seperated by Cut Quality", x = "Cut Quality", y = "Price (USD)", caption = "Diamonds data provided by ggplot2")

ggplot(data = diamonds, mapping = aes(x = reorder(cut, price, FUN = median), y = price)) +
  geom_boxplot(outlier.alpha = 0.1) + 
  labs(title = "Box Plot of Diamond Price Seperated by Cut Quality", x = "Cut Quality", y = "Price (USD)", caption = "Diamonds data provided by ggplot2")

ggplot(data = diamonds, mapping = aes(x = reorder(cut, price, FUN = median), y = price)) +
  geom_boxplot(outlier.alpha = 0.1) +
  coord_flip() + 
  labs(title = "Box Plot of Diamond Price Seperated by Cut Quality", x = "Cut Quality", y = "Price (USD)", caption = "Diamonds data provided by ggplot2")

ggplot(data = diamonds, mapping = aes(x = cut, y = price, fill = clarity)) +
  geom_boxplot(outlier.alpha = 0.1) +
  labs(title = "Box Plot of Diamond Price Seperated by Cut Quality", subtitle = "Broken into Diamond Clarity for Each Cut", x = "Cut Quality", y = "Price (USD)", legend = "Diamond Clarity", caption = "Diamonds data provided by ggplot2")

ggplot(data = diamonds, mapping = aes(x = cut, y = price)) +
  geom_point(position = "jitter", alpha = 0.1, color = "darkorange") +
  geom_boxplot(color = "blue", outlier.alpha = 0.1) + 
  labs(title = "Box Plot of Diamond Price Seperated by Cut Quality", subtitle = "Overlaid over original data to show density", x = "Cut Quality", y = "Price (USD)", caption = "Diamonds data provided by ggplot2")
```

#### Assignment 3

3. Visualize the covariation of the price of the diamonds in relation to the cut variable by using the violin plot(`geom_violin`).
    a. Mark the following quartiles in the above violin plot (0.25,0.5,0.75).

```{r}
ggplot(data = diamonds, mapping = aes(x = cut, y = price)) +
  geom_violin(draw_quantiles = c(0.25,0.5,0.75)) + 
  labs(title = "Violin Plot of Diamond Price Seperated by Cut Quality", subtitle = "Horizontal Lines Mark 25th, 50th, and 75th Percentiles", x = "Cut Quality", y = "Price (USD)", caption = "Diamonds data provided by ggplot2")
```

#### Assignment 4

4. Explore the covariation of the price and carat of diamonds using rectangular (`geom_bin2d`) and hexagonal bins (`geom_hex`).Make sure you explore this relationship with variable binwidths.

```{r}
ggplot(data = diamonds, mapping = aes(x = carat, y = price)) +
  geom_bin2d(bins = 75) +
  labs(title = "Plot of Diamond Weight vs. Price", subtitle = "Grouped into bins to show frequency", x = "Diamond Weight (carats)", y = "Price (USD)", caption = "Diamonds data provided by ggplot2")

ggplot(data = diamonds, mapping = aes(x = carat, y = price)) +
  geom_hex(bins = 75) +
  labs(title = "Plot of Diamond Weight vs. Price", subtitle = "Grouped into hexagonal bins to show frequency", x = "Diamond Weight (carats)", y = "Price (USD)", caption = "Diamonds data provided by ggplot2")
```
