---
title: "Data structures Lab Assignment"
output:
  html_document:
    df_print: paged
date: "January 31,2020"
---

Assignments:

1. Create a vector of integers, doubles, logicals and characters of your choice.
2. Create a atomic vector of numeric type and convert this vector into a matrix of your choice. 
3. Create a list and a data frame of your choice.

#### Assignment 1
1. Create a vector of integers, doubles, logicals and characters of your choice.

```{r}
int <- c(1L:5L)
int
typeof(int)
double <- 1:15
double
typeof(double)
logical <- c(T,F,T,F,F,F,T)
logical
typeof(logical)
character
character <- c("Jonathan","Carlyle","Bonaguro")
typeof(character)
```

#### Assignment 2 
2. Create a atomic vector of numeric type and convert this vector into a matrix of your choice. 

```{r}
vector_100 <- 1:100
matrix_100 <- matrix(vector_100, nrow = 10, byrow = T)
matrix_100
```

#### Assignment 3 
3. Create a list and a data frame of your choice.

```{r}
list <- list(c("Jonathan","Carlyle","Bonaguro"), 1:10, c("Python","Java","C++"))
list
dataframe <- data.frame(course_subject=c("PSYC","STAT", "PS", "PS"), instructor=c("Dr. Babu Adhimoolam", "Ha Nyguen", "Alyssa Prorok", "Aleksander Ksiazkiewicz"), course_num=c(496, 385, 396, 328), stringsAsFactors = F)
dataframe
```

