---
title: "Lab 2"
author: "Jonathan Bonaguro"
output: prettydoc::html_pretty
date: "February 5, 2020"
---

## Excercise 1

Load the data:

```{r}
deck <- read.csv("https://nkha149.github.io/stat385-sp2020/files/data/cards.csv")
```

#### Modifying the deck

* We will modify the deck for a game called “King of Hearts” where:
    + All kings get the highest value, 14 in the deck.
    + All cards in the hearts suit get the value of 10 (except the king of hearts, its value should be 14).
    
```{r}
deck[deck$suit == "hearts", 3] <- 10
deck[deck$face == "king", 3] <- 14
```

Use the following code to check your new deck:

```{r}
deck[deck$suit == "hearts", ]
deck[deck$face == "king", ]
```

#### Adding the jackpot column

* Now, create a new column called jackpot in the data frame deck.
    + Jackpot takes value of TRUE and FALSE.
    + For the king of hearts, the value of jackpot is TRUE.
    + For all other rows, the value of jackpot is FALSE.
    
```{r}
deck$jackpot <- FALSE
deck[deck$face == "king" & deck$suit =='hearts', 4] <- TRUE
deck
```

#### Function `draw()`

* Write a function draw() that randomly draws 4 cards from the deck WITHOUT replacement and compute the sum of the 4 cards that were drawn.
    + Input: no input provided.
    + Output: a number representing the sum of values of the 4 cards drawn.
    
```{r}
draw <- function(){
  four <- sample(deck$value, size = 4, replace = FALSE)
  sum(four)
}
set.seed(385)
draw()
set.seed(420)
draw()
set.seed(400)
draw()
set.seed(2020)
draw()
```

####Function `win_jackpot()`

* This is a strange game where a person can win a jackpot through 3 possible ways:
    + Getting a king of hearts in the 4 cards drawn.
    + OR getting 4 cards from the same suit.
    + OR getting 4 cards with the same face.

* Write a function `win_jackpot()` that:
    + Randomly draws 4 cards from the deck WITHOUT replacement
    + Print out the 4 cards using the print() function
    + And returns TRUE if the 4 cards drawn satisfying one of the above conditions, FALSE otherwise.
    
Hint: You might want to take a look at the unique() function to test the 2nd and 3rd conditions.

```{r}
win_jackpot <- function(){
  samp <- deck[sample(nrow(deck), 4), ]
  print(samp)
  if (TRUE %in% samp$jackpot){
    TRUE
  } else if (length(unique(samp$face)) == 1){
    TRUE
  } else if (length(unique(samp$suit)) == 1){
    TRUE
  } else {
    FALSE
  }
}
```

```{r}
set.seed(35)
win_jackpot()
set.seed(174)
win_jackpot()
```

#### Running simulation on win_jackpot()

Now, we want to estimate the probability of winning the jackpot in “King of Hearts”. How do we do that? We can do so using simulation studies, that is running the `win_jackpot()` function many many times and record the number of times we win. In this example, the number of simulations is 10000.

**<!!!>BEWARE<!!!>**: You need to change your `win_jackpot()` function before progressing to the next part.

First, write a new function called `win_jackpot2()`:
Copy the code from your original `win_jackpot()` function.
Remove the `print()` statement from the code.
Test your `win_jackpot2()` function:

```{r}
win_jackpot2 <- function(){
  samp <- deck[sample(nrow(deck), 4), ]
  if (TRUE %in% samp$jackpot){
    TRUE
  } else if (length(unique(samp$face)) == 1){
    TRUE
  } else if (length(unique(samp$suit)) == 1){
    TRUE
  } else {
    FALSE
  }
}
```

* Use the following code to run the simulation studies:
    + Make sure you copy and paste! If you retype the code, remember win_jackpot2()!!! Missing the 2 will cost     you heavily!

```{r}
results <- replicate(win_jackpot2(), n = 5000)
sum(results, na.rm = T) / 5000
```

## Exercise 2: 2015 Flight Delays and Cancellations Data

Next, we take a quick look at the 2015 Flight Delays and Cancellations Data provided by the U.S. Department of Transportation. This is a huge dataset availalbe on Kaggle. But for us, we will only take a look at flights flying out from O’Hare International Airport (ORD) in January, 2015.

```{r}
flights <- read.csv("https://nkha149.github.io/stat385-sp2020/files/data/ohare_jan.csv")
```

```{r}
nrow(flights[flights$ORIGIN_AIRPORT=="ORD",])
nrow(flights[flights$ORIGIN_AIRPORT=="ORD" & flights$AIRLINE=="AA",])
table(flights$AIRLINE)
```
