---
title: "Chapter 11: Strings with `stringr`"
author: "Jonathan Bonaguro"
output: html_document
---

```{r setup, include=FALSE, message=FALSE}
library(tidyverse)
knitr::opts_chunk$set(collapse = TRUE)
```


## Excercises: String Basics

1. In code that doesn’t use stringr, you’ll often see `paste()` and `paste0()`. What’s the difference between the two functions? What stringr function are they equivalent to? How do the functions differ in their handling of NA?

**Answer:** `paste()` is used to concatenate vectors after converting them to character, seperating them with a value specified by `sep = " "` with the default value being a space. `collapse = NULL` allows for a user to specify what character seperates the results. `paste0()` does not allow for a seperating character, and automatically smashes the strings together. However, `paste0()` does take a `collapse = NULL` argument. `str_c()` is the `stringr` function which accomplishes the same thing. It defaults to having no character to seperate the values similar to `paste0()`. `str_c()` treats missing values as true missing values, and will return `NA` similar to how `sum()` would if given an NA value. `paste()` and `paste0()` automatically convert `NA` to a string.  

2. In your own words, describe the difference between the sep and collapse arguments to str_c().

**Answer:** When combining strings into one string using `str_c()`, `sep = ""` is used to combine and seperate strings that are passed as arguments to `str_c()`. `collapse = NULL` is used when combining a vector of strings into a single string, and tells `str_c()` which values to use to seperate the items in the vector in the single string. 

3. Use `str_length()` and `str_sub()` to extract the middle character from a string. What will you do if the string has an even number of characters?

**Answer:** If the string has an even number of characters, I chose to pick extract the letter from the middle pair that is closest to the end of the word by rounding up the length divided by two using `ceiling()`.

```{r basics question 3}
x <- fruit[c(1:4, 6:8)]
x

L <-  str_length(x)
M <- ceiling(L/2)

str_sub(x, M, M)
```

4. What does `str_wrap()` do? When might you want to use it?

**Answer:** The function `str_wrap()` wraps text so that it fits within a certain width. This is useful for wrapping long strings of text to be typeset.

5. What does `str_trim()` do? What’s the opposite of `str_trim()`?

**Answer:** `str_trim()` removes whitespace from the start and end of a string. The opposite of `str_trim()` is `str_pad()`, which allows you to add whitespace to the start or end of a string. 

6. Write a function that turns (e.g.) a vector `c("a", "b", "c")` into the string `a, b, and c`. Think carefully about what it should do if given a vector of length 0, 1, or 2.

```{r basics question 6}
oxford_comma <- function(input, delim = ",") {
  n <- length(input)
  if (n == 0) {
    ""
  } else if (n == 1) {
    input
  } else if (n == 2) {
    str_c(input, collapse = " and ")
  } else {
    # commas after all n - 1 elements
    not_last <- str_c(input[seq_len(n - 1)], delim)
    # prepend "and" to the last element
    last <- str_c("and", input[[n]], sep = " ")
    # combine parts with spaces
    str_c(c(not_last, last), collapse = " ")
  }
}

oxford_comma(c("a", "b", "c", "d"))
```

## Excercises : Matching Patterns with Regular Expressions

#### Basic Matches

1. Explain why each of these strings don’t match a \: "\", "\\", "\\\".

**Answer:** A single `\` escapes the next character in an R *string*. `\\` resolves to a single `\` which escapes the next character in the *expression*. A triple `\\\`, since the doulbe slash escapes a character, will escape the escaped character. 

2. How would you match the sequence "'\?

**Answer:** The sequence `"'\` could be matched with `"\"'\\\\"`

3. What patterns will the regular expression \..\..\.. match? How would you represent it as a string?

**Answer:** This pattern will match a `.` followed by any character three times in a row, such as `.a.b.c`.

#### Anchors

1. How would you match the literal string "\$^\$"?

**Answer:** The string `$^$` could be matched with `"\\$\\^\\$"`

```{r anchors 1}
str_subset("$^$", "\\$\\^\\$")
```

2. Given the corpus of common words in stringr::words, create regular expressions that find all words that:
    a. Start with “y”.
    b. End with “x”.
    c. Are exactly three letters long. (Don’t cheat by using str_length()!)
    d. Have seven letters or more.

```{r anchors 2}
# a. Start with “y”.
str_subset(words, pattern = "^y")

# b. End with “x”.
str_subset(words, pattern = "x$")

# c. Are exactly three letters long. (Don’t cheat by using str_length()!)
str_subset(words, pattern = "^...$")

# d. Have seven letters or more.
str_subset(words, pattern = "^.......")
```
    
#### Character Classes and Alternatives

1. Create regular expressions to find all words that:
    a. Start with a vowel.
    b. Only contain consonants. (Hint: think about matching “not”-vowels.)
    c. End with ed, but not with eed.
    d. End with ing or ize.
    
```{r classes 1}
# a. Start with a vowel.
str_subset(words, pattern = "^[aeiou]")

# b. Only contain consonants. (Hint: think about matching “not”-vowels.)
str_subset(words, pattern = "^[^aeiou]+$")

# c. End with ed, but not with eed.
str_subset(words, pattern = "[^e]ed$")

# d. End with ing or ize.
str_subset(words, pattern = "i(ng|ze)$")
```

2. Empirically verify the rule “i before e except after c.”

```{r classes 2}
str_subset(words, pattern = "cei|[^c]ie")

str_subset(words, pattern = "cie|[^c]ei")
```

3. Is “q” always followed by a “u”?

```{r classes 3}
str_subset(words, pattern = "q[^u]")
```

**Answer:** Yes!

4. Write a regular expression that matches a word if it’s probably written in British English, not American English.

**Answer:** Common differences in British and American English include things like the following: `ou|ise$|ae|oe|yse$`.

5. Create a regular expression that will match telephone numbers as commonly written in your country.

**Answer:** For a format like 123-456-7891, `\\d\\d\\d-\\d\\d\\d-\\d\\d\\d\\d` would match phone numbers. 

#### Repetition

1. Describe the equivalents of ?, +, and * in {m,n} form.

**Answer:** The equivalent of `?` (0 or 1 matches) would be `{0, 1}`, the equivalent for `+` (1 or more matches) would be `{1, }`, and the equivalent for `*` (0 or more matches) would be `{0, }`.

2. Describe in words what these regular expressions match (read carefully to see if I’m using a regular expression or a string that defines a regular expression):
    a. `^.*$`
    b. `"\\{.+\\}"`
    c. `\d{4}-\d{2}-\d{2}`
    d. `"\\\\{4}"`

**Answer:** The first expression `^.*$` matches any string that starts with and has 0 or more of any character (so basically any string). The second expression `"\\{.+\\}"` matches strings that have curly braces around at least one character, e.g. `{ab}`. The third expression `\d{4}-\d{2}-\d{2}` matches 4 digits, hyphen, 2 digits, hyphen, 2 digits, similar to a date in YYYY-MM-DD form. The pattern `"\\\\{4}"` will match `\\\\` in a string. 

3. Create regular expressions to find all words that:
    a. Start with three consonants.
    b. Have three or more vowels in a row.
    c. Have two or more vowel-consonant pairs in a row.
    
```{r repetition 3}
# a. Start with three consonants.
str_subset(words, pattern = "^[^aeiou]{3}")

# b. Have three or more vowels in a row.
str_subset(words, pattern = "[aeiou]{3,}")

# c. Have two or more vowel-consonant pairs in a row.
str_subset(words, pattern = "([aeiou][^aeiou]){2,}")
```

4. Solve the beginner regexp crosswords at https://regexcrossword.com/challenges/beginner.

#### Grouping and Backreferences

1. Describe, in words, what these expressions will match:
    a. `(.)\1\1`
    b. `"(.)(.)\\2\\1"`
    c. `(..)\1`
    d. `"(.).\\1.\\1"`
    e. `"(.)(.)(.).*\\3\\2\\1"`

**Answer:** `(.)\1\1` will match any letter repeated three times, such as `"aaa"`. `"(.)(.)\\2\\1"` will match a pattern like `"abba"`. `(..)\1` will match any two characters repeated, such as `"a1a1"`. `"(.).\\1.\\1"` will match a pattern such as `"abaca"`, where the 1st, 3rd, and 5th digits are the same. `"(.)(.)(.).*\\3\\2\\1"` will match any string that has a pattern of three characters, anything in the middle, and then the three starting characters in reverse such as `"abc132cba"`.

2. Construct regular expressions to match words that:
    a. Start and end with the same character.
    b. Contain a repeated pair of letters (e.g., “church” contains “ch” repeated twice).
    c. Contain one letter repeated in at least three places (e.g., “eleven” contains three “e”s).
    
```{r grouping 2}
# a. Start and end with the same character.
str_subset(words, pattern = "^(.).*\\1$")

# b. Contain a repeated pair of letters (e.g., “church” contains “ch” repeated twice).
str_subset(words, pattern = "(..).*\\1")

# c. Contain one letter repeated in at least three places (e.g., “eleven” contains three “e”s).
str_subset(words, pattern = "(.).*\\1.*\\1")
```

## Excercises: Tools

#### Detect Matches

1. For each of the following challenges, try solving it by using both a single regular expression, and a combination of multiple str_detect() calls:
    a. Find all words that start or end with x.
    b. Find all words that start with a vowel and end with a consonant.
    c. Are there any words that contain at least one of each different vowel?
    d. What word has the highest number of vowels? What word has the highest proportion of vowels? (Hint: what is the denominator?)
    
```{r detect 1}
# a. Find all words that start or end with x.
str_subset(words, pattern = "^x|x$")
words[str_detect(words, pattern = "^x") | str_detect(words, pattern = "x$")]

# b. Find all words that start with a vowel and end with a consonant.
str_subset(words, pattern = "^[aeiou].*[^aeiou]$")

# c. Are there any words that contain at least one of each different vowel?
str_subset(words, pattern = "^[aeiou].*[^aeiou]$")

# d. What word has the highest number of vowels? What word has the highest proportion of vowels? (Hint: what is the denominator?)
```

