---
title: "Week 7 Lab Assignment 2"
output: 
  html_document:
    df_print: paged
author: "Jonathan Bonaguro"
date: "March 6, 2020"
---

```{r environment setup, echo = FALSE, message = FALSE}
library(tidyverse)
library(nycflights13)
mtcars <- as_tibble(mtcars)
band_members <- as_tibble(band_members)
band_instruments <- as_tibble(band_instruments)
```

## Assignments

1. Load the `band_members` and `band_instruments` dataset. Consider x as band_members and y as band_instruments and perform the following operations on these datasets (Make sure you use a specific key for joining the two datasets).
    a. inner_join
    b. left_join
    c. right_join
    d. full_join
    e. semi_join
    d. anti_join

2. Let `x` be `mtcars[1:18, ]` and `y` be `mtcars[12:32, ]` evaluate the following:
    a. `intersect(x,y)`
    b. `union(x,y)`
    c. `setdiff(x,y)`  

3. Join the weather and flights data sets to answer the following questions:
    a. What is the highest windspeed in the airports with maximum departure delays?
    b. What is the maximum visibiity in the airports with maximum arrival delays?

4. Join the airports and the flights datasets to answer the following questions:
    a. Can you add the location information of the airports(lat,lon,alt) to the flights dataset?
    b. What is the altitude of top 15 destinations in the year 2013?

5. Join the planes and flights dataset to answer the following questions:
    a. What is the maximum departure delay of flights manufactured before 1990?
    b. What is the maximum number of destinations travelled by planes manufactured after 2010?
    c. What is the minimum number of seats in planes travelling to top 10 destinations?

6. Using the flights dataset, find all the flights that travelled to top 3 destinations.

#### Assignment 1

1. Load the `band_members` and `band_instruments` dataset. Consider x as band_members and y as band_instruments and perform the following operations on these datasets (Make sure you use a specific key for joining the two datasets).
    a. inner_join
    b. left_join
    c. right_join
    d. full_join
    e. semi_join
    d. anti_join

```{r Assignment 1}
band_members %>% 
  inner_join(band_instruments, by = "name")

band_members %>% 
  left_join(band_instruments, by = "name")

band_members %>% 
  right_join(band_instruments, by = "name")

band_members %>% 
  full_join(band_instruments, by = "name")

band_members %>% 
  semi_join(band_instruments, by = "name")

band_members %>% 
  anti_join(band_instruments, by = "name")
```

#### Assignment 2

2. Let `x` be `mtcars[1:18, ]` and `y` be `mtcars[12:32, ]` evaluate the following:
    a. `intersect(x,y)`
    b. `union(x,y)`
    c. `setdiff(x,y)`  
    
```{r Assignment 2}
intersect(mtcars[1:18, ], mtcars[12:32, ])

union(mtcars[1:18, ], mtcars[12:32, ])

setdiff(mtcars[1:18, ], mtcars[12:32, ])
```

#### Assignment 3

3. Join the weather and flights data sets to answer the following questions:
    a. What is the highest windspeed in the airports with maximum departure delays?
    b. What is the maximum visibiity in the airports with maximum arrival delays?
    
```{r Assignment 3}
flights_weather <- flights %>%
  left_join(weather, by = c("year", "month", "day", "origin", "hour"))

flights_weather %>%
  group_by(dest) %>%
  summarize(max_dep_delay = max(dep_delay, na.rm=TRUE),
            wind_speed = max(wind_speed, na.rm = TRUE)) %>% 
  top_n(5, max_dep_delay) %>% 
  arrange(desc(max_dep_delay))

flights_weather %>%
  group_by(dest) %>%
  summarize(max_arr_delay = max(arr_delay, na.rm=TRUE),
            visib = max(visib, na.rm = TRUE)) %>% 
  top_n(5, max_arr_delay) %>% 
  arrange(desc(max_arr_delay))
  
```



#### Assignment 4

4. Join the airports and the flights datasets to answer the following questions:
    a. Can you add the location information of the airports(lat,lon,alt) to the flights dataset?
    b. What is the altitude of top 15 destinations in the year 2013?
    
```{r Assignment 4}
airport_flights <- left_join(flights, airports, by = c("dest" = "faa"))

airport_flights %>%
  group_by(dest) %>% 
  mutate(top_dest = n()) %>% 
  summarize(altitutde = max(alt, na.rm = TRUE),
            trips_to = max(top_dest)) %>% 
  top_n(15, trips_to) %>% 
  arrange(desc(trips_to))
```

#### Assignment 5

5. Join the planes and flights dataset to answer the following questions:
    a. What is the maximum departure delay of flights manufactured before 1990?
    b. What is the maximum number of destinations travelled by planes manufactured after 2010?
    c. What is the minimum number of seats in planes travelling to top 10 destinations?
    
```{r Assignment 5, message=FALSE}
planes_flights <- left_join(flights, planes, by = "tailnum")

planes_flights %>% 
  group_by(tailnum) %>% 
  filter(year.y < 1990) %>% 
  summarize(max_dep_delay = max(dep_delay, na.rm = TRUE)) %>% 
  top_n(1, max_dep_delay)

planes_flights %>% 
  group_by(tailnum) %>%
  filter(year.y > 2010) %>% 
  summarise(Unique_Dest = n_distinct(dest)) %>% 
  top_n(1, Unique_Dest)

top_dest <- flights %>%
  count(dest, sort = TRUE) %>%
  head(10)

planes_flights %>%
  semi_join(top_dest) %>% 
  filter(seats == min(seats, na.rm = TRUE)) %>% 
  summarise(seats = min(seats, na.rm = TRUE))
```

#### Assignment 6

6. Using the flights dataset, find all the flights that travelled to top 3 destinations.

```{r Assignment 6, message=FALSE}
top_dest <- flights %>%
  count(dest, sort = TRUE) %>%
  head(3)

planes_flights %>%
  semi_join(top_dest)
```

