---
title: "Week 7 Lab Assignment 1"
output: 
  html_document:
    df_print: paged
author: "Jonathan Bonaguro"
date: "March 4, 2020"
---

```{r, echo = FALSE, message = FALSE}
library(tidyverse)
library(nycflights13)
iris <- as_tibble(iris)
mtcars <- as_tibble(mtcars)
```

## Assignments

You can refer to the “Basic single-table verbs” section of the reference [dplyr page](https://dplyr.tidyverse.org/reference/index.html) for the following exercises.

**Using the flights dataset answer the following questions:**

1. Find all the flights that departed New york city in the month of January.

2. Find all the flights that departed Newyork city on January, 28th after 6 pm.

3. Find all the flights that travelled distances of greater than 3000 miles with the departure delay less than an hour.

4. Find all the flights with an arrival delay of less than an hour, but with departure delay of more than an hour.

**Using the iris dataset answer the following questions:**

5. Select the columns that starts with the characters “se”.

6. Select the columns that ends with the characters “th”.

7. Rename the species column to a column name of your choice.

**Using the mtcars dataset answer the following question.**

8. Rearrange the mtcars dataframe with values of milles per galon(mpg) in ascending order and descending order.

9. a.Group the mtcars dataset according to the mode of transmission and calculate the mean miles per gallon for each group.

9. b.Now repeat the above assignment with the pipe operator.

10. a.Group the mtcars dataset according to the number of cylinders and select the cars that have miles per galon greater than the mean mpg for each group.

10. b.Now repeat the above assignment with the pipe operator.

**Use the flights dataset for the following questions:**

11. Count the number of flights by destination.

12. Calculate the speed of the flights (speed = distance/airtime) and calculate the mean speed across different airlines.

#### Assignment 1

1. Find all the flights that departed New york city in the month of January.

```{r}
flights %>% 
  filter(month == 1)
```

#### Assignment 2

2. Find all the flights that departed Newyork city on January, 28th after 6 pm.

```{r}
flights %>% 
  filter(month == 1, day == 28, dep_time > 1800)
```

#### Assignment 3

3. Find all the flights that travelled distances of greater than 3000 miles with the departure delay less than an hour.

```{r}
flights %>% 
  filter(distance > 3000, dep_delay < 60)
```

#### Assignment 4

4. Find all the flights with an arrival delay of less than an hour, but with departure delay of more than an hour.

```{r}
flights %>% 
  filter(arr_delay < 60, dep_delay > 60)
```

#### Assignment 5

5. Select the columns that starts with the characters “se”.

```{r}
iris %>% 
  select(starts_with("se"))
```

#### Assignment 6

6. Select the columns that ends with the characters “th”.

```{r}
iris %>% 
  select(ends_with("th"))
```

#### Assignment 7

7. Rename the species column to a column name of your choice.

```{r}
iris %>% 
  rename(Breed = Species)
```

#### Assignment 8

8. Rearrange the mtcars dataframe with values of miles per gallon (mpg) in ascending order and descending order.

```{r}
mtcars %>% 
  arrange(mpg)

mtcars %>% 
  arrange(desc(mpg))
```

#### Assignment 9

9. Group the mtcars dataset according to the mode of transmission and calculate the mean miles per gallon for each group.
    b. Now repeat the above assignment with the pipe operator.

```{r}
mtcars_transmission_group <- group_by(mtcars, am)
summarize(mtcars_transmission_group, mean_mpg = mean(mpg,na.rm=TRUE))

mtcars %>%
  group_by(am) %>% 
  summarize(mean_mpg = mean(mpg, na.rm = TRUE))
```

#### Assignment 10

10. Group the mtcars dataset according to the number of cylinders and select the cars that have miles per galon greater than the mean mpg for each group.
    b. Now repeat the above assignment with the pipe operator.

```{r}
mtcars_cylinder_group <- group_by(mtcars, cyl)
filter(mtcars_cylinder_group, mpg > mean(mpg, na.rm = TRUE))

mtcars %>%
  group_by(cyl) %>% 
  filter(mpg > mean(mpg, na.rm = TRUE))
```

#### Assignment 11

11. Count the number of flights by destination.

```{r}
flights %>% 
  group_by(dest) %>% 
  summarize(number_of_flights = n())
```

#### Assignment 12

12. Calculate the speed of the flights (speed = distance/airtime) and calculate the mean speed across different airlines.

```{r}
# since air_time is in minutes, I decided to divide it by 60 to report speed in mph

flights %>% 
  mutate(speed = distance / (air_time / 60)) %>% 
  group_by(carrier) %>% 
  summarize(mean_speed = mean(speed, na.rm = TRUE))
```

