---
title: "STAT385 Notes Week 7"
author: "Jonathan Bonaguro"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(collapse = TRUE)
library(tidyverse)
```

## **Monday, Mar. 23, 2020- The Grammar of Graphics**

## `ggplot2`

* `ggplot2` is a package in the `tidyverse`
* Visualizing data using `ggplot2` has become the default for many R users
* It is much more elegant and versatile than base R graphing functions
* But that comes with the price of more "complex" function calls
    + `ggplot2` implements the grammer of graphics, a coherent system for describing and building graphs
    
## First Steps

* `ggplot2` is already installed as part of `tidyverse`
* Once you call `library(tidyverse)`, you are ready to graph using `ggplot2`
* Let's use our first graph to answer this question
    + *Do cars with big engines use more fuel than cars with small engines?*
    
### The `mpg` data frame

* To answer the question above, use the `mpg` dataset that is included in `ggplot2`

```{r}
mpg
```

* The variables we care about are
    + `displ`: a car's engine size in liters
    + `hwy`: a car's fuel efficiency on the highway in miles per gallon
    
## Our first `ggplot2` plot

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy))
```

* The above code graphs a scatterplot with `displ` on the x-axis and `hwy` on the y-axis
* Now, let's interpret the code:
    + With `ggplot2`, you begin a plot with the function `ggplot()`.
    + `ggplot()` creates a coordinate system that you can add layers to.
    + The first argument of `ggplot()` is the dataset to use in the graph.
    + So `ggplot(data = mpg)` creates an empty graph (see below)
    
```{r}
ggplot(data = mpg)
```
    
## A graphing template

* The following code segment is a reusable template for making graphs with `ggplot2`

```{r, eval=FALSE}
ggplot(data = <DATA>)
  <GEOM_FUNCTION>(mapping = aes (<MAPPINGS>))
```

## The `geom` function

* There are many `geom()` functions in ggplot2: `geom_boxplot()`, `geom_line()`, etc.
* Each `geom()` function takes a `mapping` argument. This defines how variables in your dataset are mapped to visual properties
    + The `mapping` argument is always paired with `aes()`
    + The `x` and `y` arguments of `aes()` set the axes
    + `ggplot2` looks for the mapped variables in teh data argument
    
### Examples

* **Example 1:** make a scatterplot of `hwy` vs `cyl`.
* **Example 2:** make a scatterplot of `class` vs `drv`.

```{r}
# Example 1: scatterplot of `hwy` vs `cyl`.
ggplot(data = mpg) +
  geom_point(mapping = aes(x = hwy, y = cyl))
```

```{r}
# Example 2: scatterplot of `class` vs `drv`.
ggplot(data = mpg) +
  geom_point(mapping = aes(x = class, y = drv))
```

## Aesthetic Mappings

* Aesthetic mappings use the `aes()` function
* Aesthetics are the visual properties of of the objects in your plot
* This could be the size, shape, and color of your points

### Add color to your plot

* With `aes(..., color = class)`, `ggplot2` will automatically assign a unique level of the aesthetic (here a unique color) to each unique value of the variable (here `class`), a process known as **scaling**.
* `ggplot2` will also add a legend that explains which levels correspond to which values.

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = class))
```

* Now, what’s if you don’t want to color the points by class, you just want all the points to be blue!
    + Let’s try changing class to "blue".

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy), color = "blue")
```

* This goes outside of the `aes()` argument.
    + We are manually setting the color to blue, not mapping it to any data

### Change the shape of your points

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, shape = class))
```

* `suv` is not shown because the shape palette default can only handle 6 descrete values, and `class` has 7 values. In order to have 7 classes, we would have to manually set shapes.
* We can manually set shape similar to how we manyally set color:

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy), shape = 15)
```

### Change the transparency of points

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, alpha = class))
```

* Let's try setting `alpha` to a continuous variable: 

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, alpha = cty))
```

* We can change the color to make it a little easier to read:

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, alpha = cty), color = "red")
```

### Examples

* **Example 1:** What happens if you map the same variable to multiple aesthetics?
* **Example 2:** What does the `stroke` aesthetic do? What shapes does it work with?
* **Example 3:** What happens if you map an aesthetic to something other than a variable name, like `aes(colour = displ < 5)`? Note, you’ll also need to specify x and y.

```{r}
# Example 1
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = displ))
```

* Plotting the same variable on mutliple aesthetics plots a perfectly linear correlation!

```{r}
# Example 2
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, stroke = hwy/10), shape = 21, color = "darkorange", fill = "blue")
```

* `stroke` controls the size of the outer ring the shapes 21-24. In this example, the stroke is the orange part. 

```{r}
# Example 3
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = displ < 5))
```

* In this case, the color argument took the boolean argument `displ < 5` and applied a color mapping to the variables depending on whether or not that row met that condition. 

## Common Problems

* One common problem when creating graphics using `ggplot2` is putting the `+` in the wrong places
    + It comes at the end of the line, not the start of the next line!
