---
title: "STAT385 Notes Week 7"
author: "Jonathan Bonaguro"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Wednesday, Mar. 4, 2020

**Note: I love the `tidyverse`. It's literally my favorite.**

### `tidyverse`

* The `tidyverse` contains multiple packages, such as `tidyr`, `tibble`, `dplyr`, `ggplot2`, etc...

### Tibbles

* Tibbles **are** data frames, but they tweak some old behavoirs to make life easier
* It’s difficult to change base R without breaking existing code, so most innovation occurs in packages.
    + `tidyverse` is a “mega” package (including many smaller packages).

```{r}
tibble(
  x = 1:5,
  y = 6:10
)

data.frame(
  x = 1:5,
  y = 6:10
)
```

### Prerequisites

* Install `tidyverse`
    + You only need to do this **once**.
    
```{r, eval=FALSE}
install.packages('tidyverse')
```

* "Call" the package before using it's functions
    + Need to do this everytime you open a new R session or "quit" RStudio and open it again
    
```{r, eval=FALSE}
library(tidyvers)
```

### Creating tibbles

* To create a tibble, use `tibble()`
    + Similar to `data.frame()`
    + Uses vector recycling for vectors of length 1
    
```{r}
mytb <- tibble(
  x = 1:5,
  y = 1,
  z = x ^ 2 + y
)

mytb
```

* If you try to do this using `data.frame()`, it won't work. 
    + You would need to create `x` and `y` first, then you can add `z`. It won't remember as it goes. 
    
```{r, error=TRUE}
data.frame(
  x = 1:5,
  y = 1,
  z = x ^ 2 + y
)
```

* Almost all functions in the `tidyverse` and many new R packages produce tibbles. 
* Most other (old) R packages use regular data frames.
    + to coerce a dataframe to a tibble, use `as_tibble()`
    
```{r}
mydf <- data.frame(
  x = 1:5,
  y = 1
)

as_tibble(mydf)
```

* **Note:** functions in `tidyverse` use `_` to seperate words instead of `.`
    + `read_csv()` instead of `read.csv()`
* If you're already familiar with `data.frame()`, note that `tibble()` does much less
    + It never changes the type of input
    + It never changes the names of the variables
    + It never creates row names
    
### Tibbes vs. data frames

There are two main differences in the usage of a tibble vs. a classic `data.frame`: **printing** and **subsetting**.

#### Printing

* Tibbles have a refined print method that shows:
    + only the first 10 rows,
    + all the columns that fit on screen,
    + and each column reports its type.
    
```{r}
tibble(
  a = lubridate::now() + runif(1e3) * 86400,
  b = lubridate::today() + runif(1e3) * 30,
  c = 1:1e3,
  d = runif(1e3),
  e = sample(letters, 1e3, replace = TRUE)
)
```

* But sometimes you need more output than the default display.
    + You can explicitly `print()` the data frame and control the number of rows `n` and the `width` of the display.
    + `width = Inf` will display all columns:

```{r}
nycflights13::flights

nycflights13::flights %>% 
  print(n = 10, width = Inf)
```

* **Note:** For the above code to work, you first have to install the `nycflights13` package (which includes the `flights` dataset).
* **Another note:** from now on, when I use data frame and tibble as equivalence. I will say *base R data frame* if I specifically talk about classic data frame (not tibble).    
    
#### Subsetting

* Quick review: what does `iris[,1]` return?
    + Returns the first column of `iris` as a **vector**
    
```{r}
iris[,1]
```

* What about these?
    + `iris$Sepal.Length`
    + `iris[c("Sepal.Length", "Sepal.Width")]`

```{r}
iris$Sepal.Length
iris[c("Sepal.Length", "Sepal.Width")]
```

* It returns a **vector** and a **data frame**

##### Now change `iris` to a tibble

```{r}
iris_tibble <- as_tibble(iris)
```

* Now what do these return?

```{r}
iris_tibble[,1]
iris_tibble$Sepal.Length
iris_tibble[c("Sepal.Length", "Sepal.Width")]
```

* They all return **tibbles**.

##### Summary

* `$` always returns a vector.
* With base R data frames, `[ ]` sometimes returns a data frame, and sometimes returns a vector.
* With tibbles, `[ ]` always returns another tibble.
* An even easier way to choose specific columns
    + We will learn more on Friday

```{r}
iris_tibble %>% 
  select(Sepal.Length, Sepal.Width)
```


