---
title: "STAT385 Notes Week 5"
author: "Jonathan Bonaguro"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Monday, Feb. 17, 2020

### Note

*I already have and use Git and GitHub frequently. Typing down notes for fun and to keep myself focused in class.*

### What is Git?

* A version control system.
* Its original purpose was to help groups of developers work collaboratively on big software projects. 
* Git manages the evolution of a set of files - called a **repository** - in a sane, highly structured way.

### Using Git

* There's a lot of different ways to use Git:
    + The command line terminal
    + Through a client, like RStudio
    + GitHub Desktop
    
* Introduce yourself to Git

```{r, eval=FALSE}
git config --global user.name 'Jane-Doe'
git config --global user.email 'janedoe@illinois.edu'
git config --global -list
```

### Connect to Github

* Create a repository on GitHub
* Clone the repository to your computer
* Make a local change, commit, and push
* Confirm the change on GitHub remote repository!

### How does this work?

* Pull down the repository
* Make changes in the working directory
* Stage fixes, and commit to the working directory!

#### The Three States of Git

* **Modified**
    + You have changed the file but have not committed it to your database yet.
* **Staged**
    + You have marked a modified file in its current version to go into your next commit snapshot.
* **Committed**
    + the data is safely stored in your **local** database.

#### Three Sections of a Git Project

* **Working Directory**
    + These files are pulled out of the compressed database in the Git directory and placed on disk for you     to use or modify.
* **Staging Area**
    + A file, generally contained in your Git directory, that stores information about what will go into       your next commit.
* **Git Directory**
    + Where Git stores the metadata and object database for your project. This is the most important part      of Git.

### Basic Git Workflow

* Modify files in your working tree
* Selectively stage just those changes you want to be part of your next commit, which only add those files to the commit
* You do a `commit`, which takes the files as they are in the staging area and stores that snapshot permanently to your Git directory.
* When you’re ready to update the remote repo, `push` the commit(s).







