---
title: "STAT385 Notes Week 4"
author: "Jonathan Bonaguro"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Monday, Feb. 10, 2020

#### `for` loops

* A `for` loop repeats a chunk of code many times, **one for each element in a set of input**

```{r, eval=FALSE}
for (value in that) {
  this
}
```

* The `that` should be a **set of objects ** (often a vector of numbers or character strings)
    + `1:10`
    + `1:length(x)`
    + `1:nrow(x)`
    + `c("hello", "this", "is", "a", "vector", "of", "strings")`
* The `value` symbol in a for loop acts like an argument in a function

```{r}
for (value in c("My", "first", "for", "loop")) {
  print(value)
}
```

##### Example 

* What will the value of `a` be after the loop?

```{r}
a <- 100

for (a in 1:5) {
  print(a)
}

a
```

#### Excercise 1 

* Write a `for` loop to compute the sum of 5 numbers that are randomly selected from 1 to 100 with replacement
    + First generate 5 random numbers
    + Then write the `for` loop

```{r}
# generate the five random numbers
set.seed(385)
nums <- sample(x = 1:100, size = 5, replace = TRUE)

# write the for loop
sum <- 0

for (i in 1:length(nums)) {
  sum <- sum + nums[i]
}

# another way to do it
sum2 <- 0

for (i in nums) {
  sum2 <- sum2 + i
}

sum
sum2

# check your work
sum(nums)
```

#### Working with `for` loop

* To save output from a `for` loop, you must write the loop so that it saves it's own output as it runs
* What this means:
    + create an empty `vector` or `list` before you run the loop
    + include empty `vector` or `list` in the loop
    + then you will be able to call it outside of the loop
    
```{r}
fruits <- c("apple", "pineapple", "watermelon", "orange", "peach", "plum", "honeydew", "banana", "kiwi", "papaya", "grapes", "strawberry", "blueberry", "blackberry")

fruits_short <- c()

# for loop one
for (i in fruits) {
  if (nchar(i) <= 6) {
    fruits_short <- c(fruits_short, i)
  }
}

# for loop two
fruits_short2 <- c()
count <- 0

for (i in fruits) {
  if (nchar(i) <= 6){
    count <- count + 1
    fruits_short2[count] <- i
  }
}

fruits_short
fruits_short2
```

#### `while` loops

* A `while` loop reruns a chunk **while** a certain condition remains `TRUE`

```{r, eval = FALSE}
while (condition) {
  code
}
```

* Why might `condition` change from `TRUE` to `FALSE`?
    + The code inside the loop changes the certain values which change the result of the `condition`
    + If it does not, you can end up with an infinite loop... **avoid this**
    
##### Example 

```{r}
cash <- 20
n <- 0

while (cash > 0) {
  cash <- cash - 1
  n <- n + 1
}

n
```

#### Excercise

* Remember the King of Hearts Game from lab 2?
* It costs $1 to play this game.
* If you win a jackpot, we gain $10.
* If you start with $20, how long can you play until you run out of money?

```{r, eval=FALSE}
deck <- read.csv("https://nkha149.github.io/stat385-sp2020/files/data/cards.csv")

deck$jackpot <- FALSE
deck[deck$face == "king" & deck$suit =='hearts', 4] <- TRUE

win_jackpot <- function() {
  samp <- deck[sample(nrow(deck), 4), ]
  print(samp)
  any(deck$jackpot) | length(unique(deck$face) == 1) | length(unique(deck$suit) == 1)
}

set.seed(385)
cash <- 20
n <- 0

while (cash > 0) {
  cash <- cash - 1
  n <- n + 1
  if (win_jackpot()) {
    cash <- cash + 5
  }
}

n
```

#### `repeat` loops

* The `repeat` loops repeat a chunk of code until you tell them to stop (by hittin `esc`) or until they encounter the command `break`, which stops the loop

